\documentclass{svjour3}
\usepackage[latin1]{inputenc}
\usepackage{soul}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{url}

%\there was a highlighted version provided for reviewers
%\the line below is to diable highlight
\renewcommand\hl[1]{#1}

\begin{document}
\title{Spatio-temporal Analysis of Flood Data from South Carolina}

\author{Haigang Liu* \and
        David B.\ Hitchcock \and 
        S.\ Zahra Samadi 
}
\institute{Haigang Liu, Corresponding author \at
		   Department of Statistics,  University of South Carolina\\
		   1523 Greene St, Columbia, SC 29201 \\
           \email{haigang@email.sc.edu}         
           \and
           David B.\ Hitchcock \at
           Department of Statistics,  University of South Carolina\\
		   1523 Greene St, Columbia, SC 29201 \\
\email{hitchcock@stat.sc.edu} 
           \and
           S.\ Zahra Samadi \at
           Department of Civil and Environmental Engineering, University of South Carolina\\
           301 Main St, Columbia, SC 29208\\
           \email{samadi@cec.sc.edu}   }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle
\begin{abstract}

\hl{
To investigate the relationship between flood gage height and precipitation in South Carolina from 2012 to 2016, we built a conditional autoregressive (CAR) model using a Bayesian hierarchical framework.
This approach allows the modelling of the main spatio-temporal properties of water height dynamics over multiple locations, accounting for the effect of river network, geomorphology, and forcing rainfall.
In this respect, a proximity matrix based on watershed information was used to capture the spatial structure of gage height measurements in and around South Carolina.
The temporal structure was handled by a first-order autoregressive term in the model.
Several covariates, including the elevation of the sites and effects of seasonality, were examined, along with daily rainfall amount.
A non-normal error structure was used to account for the heavy-tailed distribution of maximum gage heights.
The proposed model captured some key features of the flood process such as seasonality and a stronger association between precipitation and flooding during summer season.
The model is able to forecast short term flood gage height which is crucial for informed emergency decision.
As a byproduct, we also developed a Python library to retrieve and handle environmental data provided by some main agencies in the United States.
This library can be of general usefulness for studies requiring rainfall, flow, and geomorphological information over specific areas of the conterminous US.\
}

\keywords{Flood \and Watershed \and CAR model \and Spatiotemporal Analysis}
\end{abstract}
\newpage

\begin{center}
	\textbf{DECLARATION}
\end{center}

\section*{1. Availability of data and materials}
The datasets used and/or analyzed during the current study are available from the corresponding author upon request.
\section*{2. Competing Interests}
The authors declare that they have no competing interests.
\section*{3. Funding}
Not applicable
\section*{4. Authors' contributions}
HL analyzed scraped and prepared the data for this research and coded the model implementation and was a major contributor in writing the manuscript.
DH contributed to the methodology for statistical modeling and model validation and was also a contributor in writing.
S provided valuable domain knowledge in hydrology modeling, which helped other authors understand the data source, and is also reflected in areas such as variable selection.

\section*{5. Acknowledgements}
We are grateful for the insightful comments and suggestions of two reviewers, which have greatly improved the article.


\newpage

\section*{List of abbreviations}
\textbf{ACF}: Autocorrelation Function \\
\textbf{AWS}: Amazon Web Services \\
\textbf{CAR}: Conditional Autoregressive (Model) \\
\textbf{CONUS}: Contiguous United States \\
\textbf{COO}: Coordinate List \\
\textbf{CSR}: Compressed Sparse Row \\
\textbf{DOK}: Dictionary of Keys \\
\textbf{HUC}: Hydrologic Unit \\ 
\textbf{MAE}: Mean Absolute Error \\
\textbf{MCMC}: Markov Chain Monte Carlo \\ 
\textbf{MSE}: Mean Squared Error \\ 
\textbf{NOAA}: National Oceanic and Atmospheric Administration\\
\textbf{NWS}: National Weather Service \\ 
\textbf{PACF}: Partial Autocorrelation Function \\
\textbf{RF}: Random Forest \\ 
\textbf{SRTM}: Shuttle Radar Topography Mission \\
\textbf{SST}: Sea Surface Temperature  \\

\newpage
\section{Introduction}\label{sec:intro}
\input{1-introduction.tex}

\section{Data Sources}\label{sec:source}
\input{2-data_sources.tex}

\section{Adjacency Matrix and Watershed}\label{sec:watershed}
\input{3-adjacency-matrix-and-watershed.tex}

\section{Model Description}\label{sec:model}
\input{4-model-description.tex}

\section{Model Fitting}\label{sec:model_fitting}
\input{5-model-fitting.tex}

\section{Model Diagnosis}\label{sec:diag}
\input{6-model-diagnosis.tex}

\section{Model Comparison}\label{sec:model_comp}
\input{7-model-selection.tex}

\section{Discussion}\label{sec:discussion}
\input{8-discussion.tex}

\appendix
\section{Appendix}\label{sec:appendix}
\input{appendix}
\end{document}

